Forecasting the power of Higher Order Weak Lensing Statistics with automatically differentiable simulations

### 2021-03-09 paper/paper.tex
change short title
Added a new section

% header settings
\shorttitle{Tomographic binning optimization}
\shortauthors{Author A et al.\ (LSST~DESC)}

@@

\section{Ray-tracing}

>>>>

% header settings
\shorttitle{Differentiable Lensing Simulations}
\shortauthors{Lanzieri et al.\ (LSST~DESC)}

@@

\section{FlowPM}

\section{Ray-tracing}
In this section, we provide a brief overview of the lensing convergence and the basic concepts of cosmic shear.
Weak gravitational lensing is a promising observational technique to infer the projected matter density distribution between an observer and a source by measuring galaxy shape correlations.

In a homogeneous FLRW Universe, the geodesic equation of Light ray trajectories induced by the density fluctuations are expressed as :
\begin{equation}\label{geod}
    \frac{d^2\textbf{x}_{\bot}(\chi)}{d\chi^2}
    = -2 \nabla_{\bot} \Phi(\textbf{x}_{\bot},\chi).
\end{equation}
If we integrate Eq.\ref{geod} over the line of sight along $\chi'$ the solution $\textbf{x}_{\bot}$ for the radial comoving distance becomes:
\begin{equation}
    \textbf{x}(\chi)=f_k(\chi)\boldsymbol{\theta}-\frac{2}{c^2} 
    \int_0^{\chi} d\chi'f_k(\chi-\chi')[\boldsymbol{\nabla}_{\bot}\Phi(\textbf{x},\chi')-\boldsymbol{\nabla}_{\bot}\Phi^{(0)}(\chi')]
\end{equation}
 where we indicate as $f_k(\chi)$ and $\chi$ the angular and radial comoving distance, respectively.
 In the absence of lensing the observer would see the distance \textbf{x} under an angle $\boldsymbol{\beta}$$=\textbf{x}(\chi)/f_k(\chi)$.

 In the presence of lensing the separation vector \textbf{x} will be seen under an angle $\boldsymbol{\alpha}$, defining as:
 \begin{equation}
     \boldsymbol{\alpha}=\frac{2}{c^2} \int_0^{\chi} d\chi'
     \frac{f_k(\chi-\chi')}{f_k(\chi)}
     [\boldsymbol{\nabla}_{\bot}\Phi(\textbf{x},\chi')-\boldsymbol{\nabla}_{\bot}\Phi^{(0)}(\chi')].
 \end{equation}
 So that, the difference between the apparent angle $\theta$ and the source position $\beta$ defines the lens equation:
 \begin{equation}
 \boldsymbol{\alpha=\theta - \beta}.
 \end{equation}
In the next section we will infer the lensing potential  $\Phi$ by solving the Poisson equation from the three dimensional density contrast $\delta$:
\begin{equation}\label{Poisson}
    \nabla_{\bot}^2 \Phi(x_{\bot},\chi)=
    \frac{3H_0^2 \Omega_m}{2c^2a(\chi)}
    \delta(x_{\bot},\chi)
\end{equation}
where we indicate as $a$ the scale factor, and  as $\Omega_m$ and $H_0$ the matter density and the Hubble parameter, respectively.

\subsection{The Born approximation}

In this work we implement the Born approximation, integrating the potential gradient along the unperturbed ray. If we consider the series expansion in powers of $\Phi$, assuming that the $\Phi$ is small and considering just the first-order approximation, the gravitational potential  can be written as:
\begin{equation}
    \Phi(f_k(\chi')\boldsymbol{\theta},\chi').
\end{equation}
We can now define the Jacobian matrix \textbf{A}$=\partial\boldsymbol{\beta}/\partial\boldsymbol{\theta}$, describing the linear mapping from the lensed image to the unlensed source:
\begin{align}
    A_{ij}=& \frac{\partial\beta_i}{\partial\theta_j}=
    \delta_{ij}-\frac{\partial\alpha_i}{\partial\theta_j} \\
     & =\delta_{ij}-\frac{2}{c^2}
    \int_0^{\chi} d\chi'
     \frac{f_k(\chi-\chi')f_k(\chi')}{f_k(\chi)}
     \frac{\partial^2}{\partial x_i \partial x_j} \Phi(f_k(\chi')\boldsymbol{\theta},\chi').
\end{align}
From the parametrization of the symmetrical matrix \textbf{A}, we can define the spin-two shear $\gamma=(\gamma_1,\gamma_2)$ and the scalar convergence field, $\kappa$. 
Considering this definition the convergence and the shear are defined as second derivative of the potential:
\begin{equation}\label{kshear}
    \kappa=\frac{1}{2}(\partial_1\partial_1+\partial_2\partial_2)\psi; \ \gamma_1=\frac{1}{2}(\partial_1\partial_1-\partial_2\partial_2)\psi; \ 
    \gamma_2=\partial_1\partial_2\psi;
\end{equation}
where we define the 2D \textit{lensing potential} $\psi$,
\begin{equation}
    \psi(\boldsymbol{\theta},\chi)=
    -\frac{2}{c^2}
    \int_0^{\chi} d\chi'
     \frac{f_k(\chi-\chi')f_k(\chi')}{f_k(\chi)}
    \Phi(f_k(\chi')\boldsymbol{\theta},\chi').
\end{equation}
The two fields $\gamma$ and $\kappa$ describe the distortion in the shape of the image, and the change in the angular size, respectively.
Considering the Eq.\ref{Poisson} and the Eq.\ref{kshear} the convergence $\kappa$ becomes:
\begin{equation}
    \kappa_{born}(\boldsymbol{\theta},\chi_s)= \frac{3H_0^2 \Omega_m}{2c^2}
    \int_0^{\chi_s} 
    \frac{d\chi'}{a(\chi')}
     \frac{f_k(\chi-\chi')}{f_k(\chi)}
     f_k(\chi')
    \delta(f_k(\chi')\boldsymbol{\theta},\chi').
\end{equation}

%\begin{equation}
%where $n(\chi)$ is a redshift independent factor of normalization from Poisson equation.
 %   n(\chi)=(AB)(\chi^2)
%\end{equation}
%where $A$ is the two dimensional mesh area in $rad^2$ per pixel and $B$ is the mean 3 dimensional particle density in the all cube.
